Evaluating the Performance and Energy Efficiency of the COSMO-ART model system

### 2014-04-25 trunk/paper_enahpc14/Sec2:COSMO-ART.tex
cancels previous commits of Cristiano as IBM Zurich is actually not going to contribute to this work
inserts related work

\subsection{Model Description}
\label{subsec:1.1}
The  model  system COSMO-ART  described  in  \citep{Vogel-2009}, is  a
regional  to  continental scale  model  coupled  online  to the  COSMO
numerical weather  prediction and climate  model \citep{Baldauf-2011}.
It  incorporates  sophisticated  modules  for  gaseous  chemistry  and
aerosol dynamics  and allows the online calculation  of reactive trace
substances and their interaction  with the atmosphere.  Detailed model
description   can  be   found   in  \citep{Bangert-2012,   Knote-2011,
  Knote-2013}.

\subsection{Model Setup}
\label{subsec:1.2}
Establishing effective energy performance benchmarking of a code under
intense development such as COSMO-ART is a challenging task because of
the  absolute necessity that  results must  be reproducible  within an
expected variance for the duration of the Exa2Green project. To define
a baseline,  it was necessary  to find a run-configuration  capable of
being  recreated in  lal subsequent  versions.  Here  we give  a brief
overview of the model setup  for our performance and energy efficiency
evaluation.

Three-dimensional simulations are performed over large parts of Europe
and the Mediterranean~Sea for April 13th 2010 in order to benefit from
around  half  day  of  sun  exposure and  therefore  ensure  a  proper
activation of the chemistry cycle.  The calculation domain corresponds
to the CORDEX-EU-44 domain and is covered by a grid of $222\times 216$
points with  a horizontal resolution of $0.22\,^{\circ}$,  i.e.,  50~km
in  both directions and  40~vertical  layers.  These  24-hour forecast
simulations  are not  preceded  by  a spin-up  phase  to build-up  the
simulated gaseous and  aerosol concentrations.  This condition entails
that at the  model initialization there is a  period of adjustment for
the model  to reach its equilibrium  state and minimize  the effect of
initial conditions for gases and aerosols on the model predictions.

The meteorological initial and  boundary conditions are obtained by the
the ECMWF  global spectral model IFS  with an update  frequency of 3h.
Boundary data  for gas-phase species are taken  from IFS-MOZART output
at 6h temporal resolution. The  model setup incorporates 34~2-D and
45~3-D  fields  to be  written  out  every hour  and  is  devoid of  data
assimilation methods.

The  considered  COSMO-ART model  system  is  configured  with a  semi
Lagrangian horizontal advection scheme with tricubic interpolation and
selective filling  diffusion option in combination  with the dynamical
core  using Runge-Kutta  time  stepping \citep{COSMO-PartI-2011}.   It
also  makes use  of  the  Kinetic PreProcessor  solver  (KPP) for  the
resolution  of atmospheric  chemistry ordinary  differential equations
\citep{Damian-2002}.   Concerning the modelling  of wet  deposition in
aerosols, the  baseline has only indirect cloud  feedbacks but does not
include  in-cloud  scavenging  (rainout)  and  below-cloud  scavenging
(washout)  yet.   Amongst  physical  parameterisations,  precipitation
formation is performed by a two-moment cloud microphysics.

>>>>

\subsection{Model Description}
\label{subsec:1.1}

The model  system \textsc{Cosmo-art} described  in \citep{Vogel-2009},
is  a  regional to  continental  scale  model  coupled online  to  the
\textsc{Cosmo}   numerical  weather   prediction  and   climate  model
\citep{Baldauf-2011}.   It   incorporates  sophisticated  modules  for
gaseous  chemistry   and  aerosol  dynamics  and   allows  the  online
calculation of  reactive trace  substances and their  interaction with
the  atmosphere.    Detailed  model   description  can  be   found  in
\citep{Bangert-2012, Knote-2011, Knote-2013}.

\subsection{Model Setup}
\label{subsec:1.2}

Establishing effective energy performance benchmarking of a code under
intense development  such as \textsc{Cosmo-art} is  a challenging task
because of  the absolute necessity  that results must  be reproducible
within  an  expected  variance  for  the  duration  of  the  Exa2Green
project.  To   define  a  baseline,   it  was  necessary  to   find  a
run-configuration  capable  of   being  recreated  in  all  subsequent
versions.  Here  we give a brief  overview of the model  setup for our
performance and energy efficiency evaluation.

Three-dimensional simulations are performed over large parts of Europe
and the Mediterranean~Sea for April 13th 2010 in order to benefit from
around  half  day  of  sun  exposure and  therefore  ensure  a  proper
activation of the chemistry cycle.  The calculation domain corresponds
to the CORDEX-EU-44 domain and is covered by a grid of $222\times 216$
points with  a horizontal resolution of  $0.22\,^{\circ}$, i.e., 50~km
in  both directions  and 40~vertical  layers.  These  24-hour forecast
simulations  are not  preceded  by  a spin-up  phase  to build-up  the
simulated gaseous and  aerosol concentrations.  This condition entails
that at the model initialization,  there is a period of adjustment for
the model  to reach its equilibrium  state and minimize  the effect of
initial conditions for gases and aerosols on the model predictions.

The meteorological initial and boundary conditions are obtained by the
the ECMWF  global spectral model IFS  with an update  frequency of 3h.
Boundary data  for gas-phase species are taken  from IFS-MOZART output
at  6h temporal resolution.   The model  setup incorporates  34~2D and
45~3D  fields to  be written  out  every hour  and is  devoid of  data
assimilation methods.

The considered  \textsc{Cosmo-art} model  system is configured  with a
semi-Lagrangian    horizontal   advection    scheme    with   tricubic
interpolation  and selective filling  diffusion option  in combination
with   the   dynamical    core   using   Runge-Kutta   time   stepping
\citep{COSMO-PartI-2011}.    It  also   makes  use   of   the  Kinetic
PreProcessor solver (KPP) for  the resolution of atmospheric chemistry
ordinary  differential equations \citep{Damian-2002}.   Concerning the
modelling  of  wet  deposition  in  aerosols, the  baseline  has  only
indirect  cloud feedbacks  but  does not  include in-cloud  scavenging
(rainout) and below-cloud  scavenging (washout) yet.  Amongst physical
parameterisations,   precipitation  formation   is   performed  by   a
two-moment cloud microphysics.

### 2014-04-25 trunk/paper_enahpc14/Sec4:Experimental-results.tex
inserts related work

We  approach  the  assessment  of  the energy  footprint  and  overall
performance    of    COSMO-ART    with    two    important    metrics:
\textit{time-to-solution} (TTS) and \textit{energy-to-solution} (ETS).
TTS refers to  the total wall clock time  of the application execution
time. ETS  is the amount of  energy spent to  achieve results.  Energy
consumption  is  assessed by  sampling  the  instantaneous peak  power
during execution which  is then averaged and multiplied  by the TTS to
determine  ETS.   Whenever  possible,  multiple production  runs  were
performed  to  illustrate the  reproducibility  of  the baseline,  and
quantify the  significant uncertainties  in the power  measurement, as
dictated by the available technology.

\subsection{Time-energy analysis of COSMO-ART}
\label{subsec:4.2}

@@

\subsection{Power-performance tracing of COSMO-ART}
\label{subsec:4.3}

>>>>

We  approach  the  assessment  of  the energy  footprint  and  overall
performance   of  \textsc{Cosmo-art}   with  two   important  metrics:
\textit{time-to-solution} (TTS) and \textit{energy-to-solution} (ETS).
TTS refers to  the total wall clock time  of the application execution
time. ETS  is the amount of  energy spent to  achieve results.  Energy
consumption  is  assessed by  sampling  the  instantaneous peak  power
during execution which  is then averaged and multiplied  by the TTS to
determine  ETS.   Whenever  possible,  multiple production  runs  were
performed  to  illustrate the  reproducibility  of  the baseline,  and
quantify the  significant uncertainties  in the power  measurement, as
dictated by the available technology.

\subsection{Time-energy analysis of \textsc{Cosmo-art}}
\label{subsec:4.2}

@@

\subsection{Power-performance tracing of \textsc{Cosmo-art}}
\label{subsec:4.3}
